---
title: "P8105_hw3_yc384"
author: 'Ying Chen (UNI: yc384)'
date: "10/5/2019"
output: github_document
---

## P8105 DS HW 3

#### This homework includes figures; the readability embedded plots (e.g. font sizes, axis labels, titles)

### Problem 0

  - created a public GitHub repo and local R Project named
    p8105\_hw3\_yc384
  - created a .Rmd file named p8105\_hw3\_yc384.Rmd that renders to
    github\_document
  - created a subdirectory to store the local data files used in
    Problems 1 and 2, and use relative paths to access these data files
  - submit this link of my repo via Courseworks:
    <https://github.com/YingCarolineChen/P8105_hw3_yc384.git>

setwd(“/Users/macbook/Documents/Statistics/P8105/HW/P8105
    HW3”)

#### We first setup our default R working enviroment

``` {r setup, include=TRUE}
library (tidyverse)
library (dplyr)
library(readxl)
library(patchwork)

rm(list = ls())

knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
options(tibble.print_min = 3)
```

### **Problem 1:** This problem uses the Instacart data.

#### 1-0 Import and tidy data for hw3-problem 1

``` {r data1}
# Load dataset for problem 1 from library P8105.datasets
library(p8105.datasets)
data("instacart")

# tidy data
instacart_tidy = 
  janitor::clean_names(instacart) %>% 
  # rename long varaibles
  rename(order_hr = order_hour_of_day, order_days = days_since_prior_order) %>%
  # reorder data
  arrange(order_number, aisle_id) %>%
  select(aisle_id, product_name, order_number, order_hr, order_days, everything())

# remove duplicate rows
instacart_tidy = distinct(instacart_tidy)
```

#### 1-1 Perform EDA

#### 1-1 Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

1. We first run a summary of the dataset "instacart" that was loaded. We can see there are 15 variables and 1384617 oberserations. There are 5 different ids: order_id, producti_id, user_id, aisle_id, department_id. Depend on the needs of the analysis, we may use different ids. 

2. We then Look at the first and last 5 rows of the data 
`r head(instacart, 1)` 

We can also see that the variables that we will be using for this problem will be aisle_id, product_name, order_number and order_hour_of_day

#### days_since_prior_order?? what is this variable??

#### * 1-2-1 How many aisles are there, and which aisles are the most items ordered from?

```{r}
 instacart_tidy %>% 
  summarize(
      n_obs = n(),
      n_aisle = n_distinct(aisle_id)
            ) 
```
**Answer:** From the above result, we can see that there are 134 aisels. And the ailes that had most items ordered from are listed below:

```{r MaxOrder}
# sum all items ordered by each aisle
total =
  instacart_tidy %>% 
  group_by(aisle_id) %>% 
  mutate(total_items = sum(order_number)) %>% 
  ungroup(aisle_id) %>% 
  filter(min_rank(desc(total_items)) < 2) %>% 
  select(aisle_id, total_items)

distinct(total)
```

#### * 1-2-2 Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with > 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r Plot_Order}
total %>% 
  group_by(aisle_id) %>% 
  filter(total_items > 10000)  
  ggplot(a, aes(x = aisle_id, y = total_items, color = aisle_id)) + 
    geom_point() + geom_line() +
    scale_x_continuous(
     breaks = c(0, 20, 40, 60, 80, 100, 120, 140), 
     labels = c("0", "20", "40","60", "80","100", "120", "140")) +
    scale_y_continuous(
     breaks = c(10000, 400000, 800000, 1200000, 1600000, 2000000, 2400000, 2800000), 
     labels = c("10K", "0.4M", "0.8M", "1.2M", "1.6M", "2MK", "2.4M", "2.8M")) +
     labs(
      title = "Total Number of Items Ordered in Each Aisle (> 10K items)",
      x = "Aisle ID",
      y = "Total Number of Items Ordered (in millions)",
      caption = "Data from p8105.datasets"
      ) 
  
```

#### * 1-2-3 Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

#### * 1-2-4 Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

### **Problem 2:** This problem uses the BRFSS data.

``` {r data2}
library(p8105.datasets)
data("BRFSS")
```

#### 2-1 First, do some data cleaning: format the data to use appropriate variable names; focus on the “Overall Health” topic, include only responses from “Excellent” to “Poor”; organize responses as a factor taking levels ordered from “Poor” to “Excellent”

#### 2-2 do or answer the following (commenting on the results of each):

#### \* 2-2-1 In 2002, which states were observed at 7 or more locations? What about in 2010?

#### \* 2-2-2 Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data\_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom\_line geometry and group aesthetic will help).

#### \* 2-2-3 Make a two-panel plot showing, for the years 2006, and 2010, distribution of data\_value for responses (“Poor” to “Excellent”) among locations in NY State.

### **Problem 3:** This problem uses the accel\_data. In this spreadsheet,

variables activity.\* are the activity counts for each minute of a
24-hour day starting at
midnight.

#### 3-1-1 Load, tidy, and otherwise wrangle the data. Dataset should include all originally observed variables and values; have useful variable names;

#### 3-1-2 Include a weekday vs weekend variable; and encode data with reasonable variable classes.

#### 3-1-3 Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

#### 3-2-1 Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

#### 3-2-2 Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.

#### 3-2-3 Describe in words any patterns or conclusions you can make based on this graph.
